---
title: | 
  | \LARGE{\bf STAT 5214G: Advanced Methods of Regression}
  | \vspace{0.2cm} \Large{\bf Homework 1} 
  | \vspace{0.5cm} \large{\bf Electronically submit your solutions to Canvas by Sunday May 30th}
  | \vspace{0.2cm} {You are required to submit your solutions using R Markdown.}
author: "Krista Mosi"
date: "5.27.2021"
output: pdf_document
number_sections: TRUE
---

***

The winequalityred.csv dataset on Canvas contains information about the red wine variants of the Portuguese "Vinho Verde" wine. We are interested in determining the best model for alcohol content (y) from the available predictors: fixed acidity, volatile acidity,  citric acid, residual sugar, chorides, free sulfur dioxide, total sulfur dioxide, density, pH, and sulphates. 

```{r,echo = FALSE}

library(readr)
library(dplyr)

Data <- read.csv("C:/Users/krist/Desktop/stat 5214/winequalityred.csv")


```



1. **[12 pts]** Use scatterplots (a matrix is fine) to assess the individual relationship of each predictor with the response. Comment on what you observe. 

```{r, echo = FALSE}
library(ggplot2)

plot1 <- ggplot(data = Data, aes(y= alcohol, x = fixed.acidity)) + geom_point(data = Data) +theme_bw()

plot2 <- ggplot(data = Data, aes(y= alcohol, x = volatile.acidity)) + geom_point(data = Data) +theme_bw()

plot3 <- ggplot(data = Data, aes(y= alcohol, x = citric.acid)) + geom_point(data = Data) +theme_bw()

plot4 <- ggplot(data = Data, aes(y= alcohol, x = residual.sugar)) + geom_point(data = Data) +theme_bw()

plot5 <- ggplot(data = Data, aes(y= alcohol, x = chlorides)) + geom_point(data = Data) +theme_bw()

plot6 <- ggplot(data = Data, aes(y= alcohol, x =free.sulfur.dioxide)) + geom_point(data = Data) +theme_bw()

plot7 <- ggplot(data = Data, aes(y= alcohol, x = total.sulfur.dioxide)) + geom_point(data = Data) +theme_bw()

plot8 <- ggplot(data = Data, aes(y= alcohol, x = density)) + geom_point(data = Data) +theme_bw()

plot9 <- ggplot(data = Data, aes(y= alcohol, x = pH)) + geom_point(data = Data) +theme_bw()

plot10 <- ggplot(data = Data, aes(y= alcohol, x = sulphates)) + geom_point(data = Data) +theme_bw()


gridExtra::grid.arrange(plot1,plot2, plot3, plot4, plot5, ncol = 2)


```

*   For the first group of plots, all 5 have the majority of their points concentrated on the left side of the plot. Chlorides and residual sugar especially.  They both show a pattern that lower levels of chlorides/residual sugars, have higher alcohol content. 

*   Citric Acid and fixed acity have a larger spread of points and more variability when it comes to their levels and alcohol content.


```{r, echo = FALSE}
gridExtra::grid.arrange(plot7,plot8, plot9, plot10, plot6, ncol = 2)
```



*   This group of plots shows a few different patterns. Most noticeably is the relationship between alcohol and density. The pattern shows higher levels of alcohol content with lower density.

*   Total sulfur dioxide and free sulfur dioxide are mainly concentrated in the left side of the graph, but free sulfur dioxide has more of a spread. 

*   PH levels show a concentration in the middle, primarily clustered between 3.0 and 3.5, and a range of alcohol content.  Sulphates has a pattern where levels are primarily between .5 and 1.0 and alcohol content primarily under 14.


2. **[12 pts]** Use	R	to	 fit	a	multiple	linear	regression	model	with	all	the	predictors.	Provide	the	R	output.
```{r, echo = FALSE}

Full_model <- lm(alcohol ~., data = Data)

print(Full_model)

summary(Full_model)
```



3. **[21 pts]** Use	 Forward	 selection,	 Backward	 elimination,	 and	 Stepwise	 regression	 procedures	 to	 find	 a	 model.	Compare	the	models	in	terms	of	AIC,	R2adj,	and included	variables.	 Select	a	wining	model	from those	obtained	here. Provide the estimate table of your winning model. 

### Forward Selection

```{r, echo = FALSE}
library(olsrr)

Forward = ols_step_forward_p(Full_model, penter = .05)

Forward

Forward_2=ols_step_forward_aic(Full_model)

Forward_2

```
### Backwards

```{r, echo = FALSE}


Backward = ols_step_backward_p(Full_model, penter = .05)

Backward

Backward_2=ols_step_backward_aic(Full_model)
Backward_2
```
### Stepwise 


```{r, echo = FALSE}
library(olsrr)

Stepwise= ols_step_both_p(Full_model,prem=0.05)

Stepwise

Stepwise_2=ols_step_both_aic(Full_model)
Stepwise_2
```

*   When comparing the models, we see that using the model without free sulfur dioxide is the best option. Forward and stepwise regression add all the variables except free sulfur dioxide into the model. This model is also found when looking at Backwards regression using AIC.

*   We can see that this  model with 9 variables (shown in Forward, Stepwise, and backwards using AIC) has the highest Adjusted r-squared value at .668, the lowest AIC value at 2990.052.

### Estimate table 

```{r, echo=FALSE}
Stepwise$model

winning_model <- lm(alcohol ~ density+ fixed.acidity + pH + residual.sugar + sulphates + citric.acid + total.sulfur.dioxide + chlorides + volatile.acidity, data = Data)

summary(winning_model)
```

4. **[12 pts]** Test whether alcohol is dependent	on	at	least	one	of	the	predictors (in your winning model). Provide	the	hypotheses,	test	statistic,	p-value	and	your	conclusion	in	terms	of	the	problem.
```{r, echo = FALSE}
winning_model <- lm(alcohol ~ density+ fixed.acidity + pH + residual.sugar + sulphates + citric.acid + total.sulfur.dioxide + chlorides + volatile.acidity, data = Data)

summary(winning_model)

```
Hypothesis: 

$$H_0 : \beta_1 = \beta_2 = \cdots = \beta_k = 0 $$

$$H_A : \beta_j \ne 0 $$, for at least one j

Conclusion: With a p-value of 2.2e-16, which is statistically significant at the .05 level, we reject the null hypothesis and conclude that at least one of our explanatory variables contributes significantly to the model.



5. **[12 pts]** Perform	 the	 hypothesis	 test	 to	 determine	 whether	 the	 alcohol content	 is	associated with	chlorides.	Provide	the	hypotheses,	test	statistic,	p-value	and	your	conclusion	in	terms	of	
the	problem.
```{r, echo = FALSE}
summary(winning_model)
```


Hypothesis:  

$$H_0 : \beta_j = 0$$

$$H_A : \beta_j \ne 0 $$

Conclusion: With a p-value of 2.2e-16, which is statistically significant at the .05 level,  we reject the null hypothesis and conclude that sulphates  contributes significantly to the model.



6. **[10 pts]** Interpret	the	estimated	regression coefficient for	sulphates.

An increase in sulphates associated with an increase in alcohol content.  If all other variables were held constant, alcohol content would increase  by 1.242 with a one unit increase of sulphates.


7. **[11 pts]** Construct	a	95%	confidence interval	for	the	regression	parameters.

I did this for both the full model and the winning model chosen:


```{r, echo = FALSE}
confint(Full_model,level=0.95)

confint(winning_model, level = .95)

```



8. **[10 pts]** Suppose you want to predict the alcohol content for a new wine the characteristics below. Construct an appropriate interval. 
    + fixed.acidity = 7.6
    + volatile.acidity = 0.35
    + citric.acid = 0.1
    + residual.sugar = 2.0
    + chlorides = 0.087
    + total.sulfur.dioxide = 59
    + density = 0.99
    + pH = 3.45
    + sulphates = 0.63
    
```{r, echo = FALSE}

NewData <- data.frame(fixed.acidity = 7.6,
    volatile.acidity = 0.35,
     citric.acid = 0.1,
     residual.sugar = 2.0,
     chlorides = 0.087,
     total.sulfur.dioxide = 59,
     density = 0.99,
     pH = 3.45,
     sulphates = 0.63 )


predict(winning_model, newdata = NewData, interval = 'prediction', level =.95)
```

    